Model Finding for Recursive Functions in SMT\thanks{ This research is partially supported by the Inria technological development action ``Contre-exemples Utilisables par Isabelle et Coq'' (CUIC). }

### 2015-05-07 tex/main.tex
more

%\section{A Translation for Admissible Function Definitions}
%\section{A Translation for Function Definitions}
\section{The Translation}
\label{sec:encoding}

\begin{figure}[t]
\normalsize
\begin{enumerate}
%\begin{framed}
\item[\ ]
$\conv_0( t\typ{\tau},\, p )$ $=$
\\[-.8\baselineskip]
 \begin{itemize}
   \item[] $\mathsf{if}$ $\tau = \Bool$ and $t = \boolop(t_1,\ldots,t_n)$ $\mathsf{then}$
    \begin{itemize}
      \item[] $\mathsf{let}$ $( t'_i,\, \chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ $\mathsf{for}$ $i = 1, \ldots, n$ $\mathsf{in}$
      \item[] $\mathsf{let}$ $\chi = \chi_1 \wedge \cdots \wedge \chi_n$ $\mathsf{in}$
      \item[] $\mathsf{if}$ $p = \ppos$ $\mathsf{then}$
      \begin{itemize}
        \item[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n) ) \wedge \chi,\, \ltrue$}
      \end{itemize}
      \item[] $\mathsf{else}$ $\mathsf{if}$ $p = \pneg$ $\mathsf{then}$
      \begin{itemize}
        \item[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n) \vee \neg \chi,\, \ltrue$}
      \end{itemize}
      \item[] $\mathsf{else}$
      \begin{itemize}
        \item[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n),\, \chi$}
\\[-.8\baselineskip]
      \end{itemize}
    \end{itemize}
  \item[] $\mathsf{else}$ $\mathsf{if}$ $t = \forallf{\con{f}} \vec x.\;\, u$ $\mathsf{then}$
    \begin{itemize}
      \item[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\mathsf{in}$
      %\item[] $\mathsf{assert}( \chi = \ltrue )$;
      \item[] \return \bigtuple{$\forall a : \fargtype{\con{f}}.\; u' [ \vecfarg{\con{f}}( a ) / \vec x ],\, \ltrue$}
\\[-.8\baselineskip]
    \end{itemize}
  \item[] $\mathsf{else}$ $\mathsf{if}$ $t = \forall \vec x.\;\, u$ $\mathsf{then}$
    \begin{itemize}
      \item[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\mathsf{in}$
      \item[] \return \bigtuple{$\forall \vec x.\; u',\, \forall \vec x.\; \chi$}
\\[-.8\baselineskip]
    \end{itemize}
   \item[] $\mathsf{else}$
   \begin{itemize}
     \item[] \return \bigtuple{$t,\; \bigwedge\, \{ \exists a : \fargtype{\con{f}}.\; \vecfarg{\con{f}}( a ) \teq \vec t \mid \con{f}( \vec t ) \in \terms( t ),\, \con{f} \in \sfundefs{\Sigma} \}$}
   \end{itemize}
 \end{itemize}
\end{enumerate}
\begin{enumerate}
%\begin{framed}
\item[\ ]
$\conv( \varphi )$ $=$
\\[-.8\baselineskip]
 \begin{itemize}
   \item[] $\mathsf{let}$ $( \varphi',\, \_ ) = \conv_0( \varphi,\, \ppos )$ $\mathsf{in}$
   %\item[] $\mathsf{assert}( D = \emptyset )$;
   \item[] \return $\varphi'$
  \end{itemize}
\end{enumerate}
\caption{\,Definition of function $\conv$}
\label{fig:A}
\end{figure}

@@

If $t$ is a quantified formula, $\conv_0$ recursively calls itself on the body
with the same polarity. If $t$ is a function definition of a function in
$\sfundefs{\Sigma}$, $\conv_0$ constructs a quantified formula over a single variable
$a$ of type $\fargtype{\con{f}}$ and replaces all occurrences of $\vec x$ in
$t'_1$ with $\vecfarg{\con{f}}( a )$. Since function definitions
are a top-level conjunct, %by case analysis on the return values of $\conv$,
we know that $\chi$ is simply $\ltrue$.
If $t$ is not a function definition, the resulting
quantified formula is reconstructed with the body $u'$ returned by the
recursive call, and a quantifier is prefixed to the constraint $\chi$
returned by the call.

>>>>

\section{The Translation}
\label{sec:encoding}

%%% TYPESETTING: approximated, to get centering
\newcommand{\itemx}{\itemindent5em\item}

\begin{figure}[t]
\normalsize
\begin{enumerate}
%\begin{framed}
\itemx[\ ]
$\conv_0( t\typ{\tau},\, p )$ $=$
\\[-.8\baselineskip]
 \begin{itemize}
   \itemx[] $\mathsf{if}$ $\tau = \Bool$ and $t = \boolop(t_1,\ldots,t_n)$ $\mathsf{then}$
    \begin{itemize}
      \itemx[] $\mathsf{let}$ $( t'_i,\, \chi_i ) = \conv_0( t_i,\, \pol( \boolop,\, i,\, p ) )$ $\mathsf{for}$ $i = 1, \ldots, n$ $\mathsf{in}$%\kern5em$\mid$
      \itemx[] $\mathsf{let}$ $\chi = \chi_1 \wedge \cdots \wedge \chi_n$ $\mathsf{in}$
      \itemx[] $\mathsf{if}$ $p = \ppos$ $\mathsf{then}$
      \begin{itemize}
        \itemx[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n) ) \wedge \chi,\, \ltrue$}
      \end{itemize}
      \itemx[] $\mathsf{else}$ $\mathsf{if}$ $p = \pneg$ $\mathsf{then}$
      \begin{itemize}
        \itemx[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n) \vee \neg \chi,\, \ltrue$}
      \end{itemize}
      \itemx[] $\mathsf{else}$
      \begin{itemize}
        \itemx[] \return \bigtuple{$\boolop(t'_1, \ldots, t'_n),\, \chi$}
\\[-.8\baselineskip]
      \end{itemize}
    \end{itemize}
  \itemx[] $\mathsf{else}$ $\mathsf{if}$ $t = \forallf{\con{f}} \vec x.\;\, u$ $\mathsf{then}$
    \begin{itemize}
      \itemx[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\mathsf{in}$
      %\itemx[] $\mathsf{assert}( \chi = \ltrue )$;
      \itemx[] \return \bigtuple{$\forall a : \fargtype{\con{f}}.\; u' [ \vecfarg{\con{f}}( a ) / \vec x ],\, \ltrue$}
\\[-.8\baselineskip]
    \end{itemize}
  \itemx[] $\mathsf{else}$ $\mathsf{if}$ $t = \forall \vec x.\;\, u$ $\mathsf{then}$
    \begin{itemize}
      \itemx[] $\mathsf{let}$ $( u',\, \chi ) = \conv_0( u,\, p )$ $\mathsf{in}$
      \itemx[] \return \bigtuple{$\forall \vec x.\; u',\, \forall \vec x.\; \chi$}
\\[-.8\baselineskip]
    \end{itemize}
   \itemx[] $\mathsf{else}$
   \begin{itemize}
     \itemx[] \return \bigtuple{$t,\; \bigwedge\, \{ \exists a : \fargtype{\con{f}}.\; \vecfarg{\con{f}}( a ) \teq \vec t \mid \con{f}( \vec t ) \in \terms( t ),\, \con{f} \in \sfundefs{\Sigma} \}$}
   \end{itemize}
 \end{itemize}
\end{enumerate}
\begin{enumerate}
%\begin{framed}
\itemx[\ ]
$\conv( \varphi )$ $=$
\\[-.8\baselineskip]
 \begin{itemize}
   \itemx[] $\mathsf{let}$ $( \varphi',\, \_ ) = \conv_0( \varphi,\, \ppos )$ $\mathsf{in}$
   %\itemx[] $\mathsf{assert}( D = \emptyset )$;
   \itemx[] \return $\varphi'$
  \end{itemize}
\end{enumerate}
\caption{\,Definition of function $\conv$}
\label{fig:A}
\end{figure}

@@

If $t$ is a function definition of a function in
$\sfundefs{\Sigma}$, $\conv_0$ constructs a quantified formula over a single variable
$a$ of type $\fargtype{\con{f}}$ and replaces all occurrences of $\vec x$ in
$t'_1$ with $\vecfarg{\con{f}}( a )$. Since function definitions
are a top-level conjunct, %by case analysis on the return values of $\conv$,
$\chi$ must be $\ltrue$.

If $t$ is an unannotated quantified formula, $\conv_0$ recursively calls itself
on the body with the same polarity. A quantifier is prefixed to the
quantified formula and constraint returned by the recursive call.
